|

### 2023-01-18 paper/paper.Rmd
fig 3 revised

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data. The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
#ggscatmat(penguins_data, columns = 2:5, col="species") +
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  theme(aspect.ratio=1, 
        #legend.position = "bottom", 
        axis.text = element_blank(),
        axis.ticks = element_blank())
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  ylim(c(0,4))
pscat
```

@@

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left). We can clearly see a block structure, while the LDA model (middle) produces linear boundaries. The three groups are nicely separated in this projection of the data (right).}
\label{proj1}
\end{figure*}

>>>>

```{r penguins-scatmat, out.width="60%", width=10, height=10, fig.align='center', fig.cap="Scatterplot matrix of the (standardised) penguins data. The three species are reasonably different in size, with Gentoo distinguised from the other two on body depth relative to flipper length and body mass."}
library(tidyverse)
library(palmerpenguins)
library(GGally)
penguins_data <- penguins[, c(1,3, 4, 5, 6)] %>%
  drop_na() %>%
  mutate(across(where(is.numeric), function(x) (x-mean(x))/sd(x))) %>%
  rename("bl"="bill_length_mm",
         "bd"="bill_depth_mm",
         "fl"="flipper_length_mm",
         "bm"="body_mass_g")
#ggscatmat(penguins_data, columns = 2:5, col="species") +
# library(paletteer) 
# rcartocolor::ag_Sunset
# rcartocolor::ag_GrnYl
# RColorBrewer::Spectral
# clrs <- c("#245668FF", "#39AB7EFF", "#A9DC67FF")
# clrs <- c("#4B2991FF", "#EA4F88FF", "#F6A97AFF")
# clrs <- c("#4B2991FF", "#39AB7EFF", "#F6A97AFF")
clrs <- c("#245668FF", "#EA4F88FF", "#ABDDA4FF")
penguins_diag <- tibble(x=1, y=c(3,2,1),
     species=levels(penguins_data$species))
pscat <- ggpairs(penguins_data, columns = 2:5, 
          mapping = ggplot2::aes(color = species, 
                                 fill = species),
          upper = list(continuous = "points"),
          diag = list(continuous = "densityDiag")) +
#  scale_colour_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
#  scale_fill_manual("", values=c("Adelie" = "#5E82B6", "Chinstrap" = "#E19C24", "Gentoo" = "#8FB132")) +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  scale_fill_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) 
pscat[1,1] <- ggplot(penguins_diag, 
                     aes(x=x, 
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[2,2] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[3,3] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat[4,4] <- ggplot(penguins_diag, 
                     aes(x=x,
                         y=y, 
                         label=species, 
                         colour=species)) + 
  geom_text() +
  scale_colour_manual("", values=c("Adelie" = clrs[1], "Chinstrap" = clrs[2], "Gentoo" = clrs[3])) +
  ylim(c(0,4))
pscat + theme_bw() +
  theme(aspect.ratio=1, 
        panel.grid.major = element_blank(),
        panel.grid.minor = element_blank(),
        axis.text = element_blank(),
        axis.ticks = element_blank())
```

@@

\begin{figure*}[ht]
\centerline{\includegraphics[width=0.95\textwidth]{figures/proj1.pdf}}
%\centerline{\includegraphics[width=0.32\textwidth]{figures/proj1_rf.png}
%\includegraphics[width=0.32\textwidth]{figures/proj1_lda.png}
%\includegraphics[width=0.32\textwidth]{figures/proj1_data.png}}
\caption{Projection identified using the manual tour, because it reveals an interesting structure in the predictions from the RF model (left). We can clearly see a block structure, while the LDA model (middle) produces linear boundaries. The three groups are nicely separated in this projection of the data (right).}
\label{proj1}
\end{figure*}
